### 2018-02-11 article.tex
Text on degrees of polynomials

\documentclass[a4paper]{article}
\usepackage[margin=1.0in]{geometry}

\usepackage[l2tabu,orthodox]{nag}
\usepackage{indentfirst}
\usepackage{amssymb,amsfonts}
\usepackage[]{mathtools}
\usepackage{cmap}
\usepackage[T2A]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{ucs}
\usepackage[russian,english]{babel}
\usepackage[babel = true]{microtype}
\usepackage{graphicx}
\usepackage[colorinlistoftodos, textsize=tiny]{todonotes}

\usepackage{color}
\definecolor{darkblue}{rgb}{0,0,.75}
\definecolor{darkred}{rgb}{.7,0,0}
\definecolor{darkgreen}{rgb}{0,.7,0}

\usepackage[
    draft = false,
    unicode = true,
    colorlinks = true,
    allcolors = blue,
    hyperfootnotes = true
]{hyperref}
\usepackage{amsmath}
\usepackage{amsthm}
    \theoremstyle{plain}
    \newtheorem{theorem}{Theorem}[section]
    \newtheorem{task}{Задача}[section]
    \newtheorem{lemma}{Lemma}[section]
    \newtheorem{proposition}{Proposition}[section]
    \newtheorem{corollary}{Corollary}[section]
    \theoremstyle{definition}
    \newtheorem{definition}{Definition}
    \newtheorem*{notation}{Notation}
    \newtheorem*{example}{Example}
        \newtheorem*{answer}{Ответ}
    \newtheorem*{draft}{Черновик ответа}
\usepackage{thmtools}

\usepackage{hyperref}
\usepackage[nameinlink]{cleveref}

\makeatletter
\newcommand*{\indep}{%
  \mathbin{%
    \mathpalette{\@indep}{}%
  }%
}
\newcommand*{\@indep}[2]{%
  % #1: math style
  % #2: empty or \not
  \sbox0{$#1\perp\m@th$}%        box 0 contains \perp symbol
  \sbox2{$#1=$}%                 box 2 for the height of =
  \sbox4{$#1\vcenter{}$}%        box 4 for the height of the math axis
  \rlap{\copy0}%                 first \perp
  \dimen@=\dimexpr\ht2-\ht4-.2pt\relax
      % The equals symbol is centered around the math axis.
      % The following equations are used to calculate the
      % right shift of the second \perp:
      % [1] ht(equals) - ht(math_axis) = line_width + 0.5 gap
      % [2] right_shift(second_perp) = line_width + gap
      % The line width is approximated by the default line width of 0.4pt
  \kern\dimen@
  {#2}%
      % {\not} in case of \nindep;
      % the braces convert the relational symbol \not to an ordinary
      % math object without additional horizontal spacing.
  \kern\dimen@
  \copy0 %                       second \perp
}

\usepackage{bbm}

\begin{document}
\section{Basic definitions}
$\Gamma := \mathbb{Z}_p = \left\{x = \left(x_0, x_1, x_2, \ldots \right), x_k \in \{0, 1, \ldots, p - 1\} \right\}$

$\Gamma' := \Gamma \setminus \{(p-1,p-1,\ldots)\}$

$\lambda$ is the Haar measure on $\Gamma$.

    We define two $\lambda$-preserving transformations on $\Gamma$:
    \begin{itemize}
    \item $\sigma:\ x=\left(x_0, x_1, x_2, \ldots \right) \mapsto \sigma x = \left(x_1, x_2, \ldots \right)$
    \item $S:\ x \mapsto x + 1$, where $1:=(1,0,0,\ldots)$
    \end{itemize}
    Let $\phi: \Gamma' \rightarrow \mathbb{Z}$ be the <<first not $(p-1)$>> functional:\\
    \[\phi(x):=x_i\text{ if }x=(p-1, \ldots, p-1, x_i, x_{i+1}, \ldots)\]
    \[\phi^{(0)}(x):=0;\ \phi^{(m)}(x):=\phi(x)+\phi(Sx)+\ldots+\phi(S^{m-1}x)\]

Let us define $\pi_m$ as the probability distribution of $\phi^{(m)}$ on $\mathbb{Z}$:
    \[\pi_m(j) = \lambda(\phi^{(m)}=j)\]
    We consider the sequences of polynomials $P_m^p$ produced by $\pi_m$ for fixed $p$ and $m$:
    \[P_m^p(t):= \mathbb{E}_\lambda\left[ t^{\phi^{(m)}(x)}\right] = \sum\limits_{j=0}^m \pi_m(j) t^j \]
\section{Palindromic property}
Let $p \ge 3$ be an \textit{odd} number.

We can generalize the definition of $\phi$ functional:
\begin{definition}
    $$
    \phi(x) = \begin{cases}
                    \omega(x_0), & 0 \le x_0 \le p - 2 \\
                    \phi(\sigma x), & x_0 = p - 1
                \end{cases}
    $$
\end{definition}
    The classic case is hence described by $\omega(j)=j$. This is the only non-trivial $\omega$ for $p=3$. \\

\begin{definition}
Let $\phi_\star = \min\limits_x \phi(x), \phi^\star = \max\limits_x \phi(x), \delta=\phi^\star - \phi_\star$. We say that a functional $\phi$  \textit{has the palindromic property} iff 
$$
\forall j,\ 0 \le j \le m\delta:   \pi_m(m\phi_\star + j)=\pi_m(m\phi^\star-j)
$$
\end{definition}

    Let us find the sufficient conditions for $\phi$ to have the palindromic property.

\begin{definition}
We call a function $\omega$  \textit{binary-antipalindromic} iff
\[\begin{cases}
	\mathrm{Ran }\ \omega = \{0,1\} \\
	\forall j,\ 0 \le j \le p-2: \  \omega(j) = 1 - \omega(p-2-j)
\end{cases}\]
\end{definition}

\begin{example}
The usual $\omega(j)=j$ is binary-antipalindromic for $p = 3$. Indeed,
$$
\omega(0) = 0 = 1 - \omega(1) = 1 - \omega(3 - 2 - 0)
$$
$$
\omega(1) = 1 = 1 - \omega(0) = 1 - \omega(3 - 2 - 1)
$$
\end{example}

\begin{theorem}
If $\omega$ is binary-antipalindromic, the corresponding $\phi$ has the palindromic property.
\end{theorem}

We will need the following lemmas to prove the theorem. (Note that the lemmas generalize Lemma 3.1 and Lemma 3.2 from the article "Weak limits of powers of Chacon's automorphism"\ )

\begin{lemma} \label{binDistLemma}
If $\omega$ is binary-antipalindromic, the probability distributions of the random sequences $\{\phi(S^j x)\}_{j \ge 0}$ and $\{ 1 - \phi(S^{-j} x)\}_{j \ge 0}$ are the same.
\end{lemma}
\begin{proof}
Let $x \in \Gamma \setminus \{(p-1,p-1,\ldots)\}$. We
say that $\mathrm{order}(x) = k \ge 0$ if $x_0=x_1=\ldots=x_{k-1}=p-1$ and $x_k \ne p-1$.
Since the first digit in the sequence $(\ldots, x-1,x,x+1)$ follows a periodic pattern
$(\ldots, 0, 1, 2, \ldots,p-2, p-1, 0, 1, 2, \ldots,p-2, p-1, \ldots)$, the contribution of points of order $0$ in the sequence
$\{\phi(S^j x)\}_{j \ge 0}$ provides a sequence of blocks $\omega(0),\omega(1),\ldots, \omega(p-2)$ separated by one symbol
given by a point of higher order. To fill in the missing symbols corresponding to positions $j$ 
such that $\mathrm{order}(x+j) \ge 1$, we observe that, if $x$ starts with a $p-1$, then for all $j \ge 0$, $\phi(x+pj)=\phi(\sigma x + j)$. Hence the missing symbols are given by the symbols $\{\phi(S^j \sigma x)\}_{j \ge 0}$.\\
Let us observe an example for $p=7$ produced by the Legendre symbol-like $\omega(x)=((\frac{x+1}{7}) + 1)/2$ (which is binary-antipalindromic due to the Quadratic reciprocity):\bigskip\\
$\begin{array}{cccccccccccccccccccccccccccl}
1 & 1 & 0 & 1 & 0 & 0 & . & 1 & 1 & 0 & 1 & 0 & 0 &. & 1 & 1 & 0 & 1 & 0 & 0 & . & 1 & 1 & 0 & 1 & 0 & 0 &\leftarrow \text{contribution of order }0 \\
  &   &   &   &   &   & 1 &   &   &   &   &   &   &1 &   &   &   &   &   &   & 0 &   &   &   &   &   &   &\leftarrow \text{contribution of order }1 \\
1 & 1 & 0 & 1 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 1 & 1 & 1 & 0 & 1 & 0 & 0 & 0 & 1 & 1 & 0 & 1 & 0 & 0 &\leftarrow \text{the whole sequence} \\
\end{array}$
\bigskip\\
We may build $\{ 1 - \phi(S^{-j} x)\}_{j \ge 0}$ from $\{\phi(S^j x)\}_{j \ge 0}$ with a composition of two measure-preserving transformations:
\begin{enumerate}
\item The 'reverse' transformation  $\{a_j\}_{j \ge 0} \mapsto \{a_{-j}\}_{j \ge 0}$ 
\item The 'flip' transformation which substitutes each element $k \in \{b_j\}_{j \ge 0}$ with $1-k$, turning the sequence into $\{ 1 - b_j\}_{j \ge 0}$ .
\end{enumerate}
By the definition of binary-antipalindromic, it is clear that this procedure works as identity transformation when applied to $\{\phi(S^j x)\}_{j \ge 0}$.
\end{proof}
\begin{lemma}
If $\{\phi(S^j x)\}_{j \ge 0} \overset{d}{=} \{ 1 - \phi(S^{-j} x)\}_{j \ge 0}$, then for any $0 \le j \le m:\ \pi_m(j)=\pi_m(m-j)$
\end{lemma}
\begin{proof}
The coefficient $\pi_m(m-j)$ is equal to the probability to see $(m-j)$ digits equal to 1 in the $m$ consecutive terms in $\{\phi(S^j x)\}_{j \ge 0}$. Thus, $\pi_m(m-j)$ is also equal to the probability to see $j$ digits equal to $0$ in the $m$ consecutive terms. Using \Cref{binDistLemma}, we conclude the statement of this lemma.\\
\end{proof}

To prove the theroem, it is now enough to notice that in Definition 1 for $\phi$ produced by binary-antipalindromic $\omega$ we may substitute $\phi_\star = 0, \phi^\star = 1, \zeta=1$ and hence the palindromic properly is equivalent to $\forall j, 0 \le j \le m:\ \pi_m(j)=\pi_m(m-j)$.
This summarizes the consideration of $\omega$ functions with $\mathrm{Ran }\ \omega = \{0,1\} $. Let us skip to the general construction.

\begin{definition}
We call a function $\omega$  \textit{antipalindromic} iff
\[\begin{cases}
	\mathrm{Ran }\ \omega = \{0,1, \ldots, \zeta\} \\
	\forall j,\ 0 \le j \le p-2: \  \omega(j) = \zeta - \omega(p-2-j)
\end{cases}\]
\end{definition}

\begin{notation}
We further use $[M, N]$ instead of $\{M,M+1, \ldots, N\}$ for shortness.
\end{notation}
Note that given $\omega$ such that $\mathrm{Ran}\ \omega = [0,\zeta]$, we have $\mathrm{Ran}\ \phi = [0,\zeta]$ and $\mathrm{Ran}\ \phi^{(m)} = [0,m\zeta]$

\begin{lemma} \label{distLemma}
If $\omega$ is antipalindromic, the probability distributions of the random sequences $\{\phi(S^j x)\}_{j \ge 0}$ and $\{ \zeta - \phi(S^{-j} x)\}_{j \ge 0}$ are the same.
\end{lemma}
This lemma is shown similarly to the \Cref{binDistLemma}.\bigskip \\
The following theorem generalizes Theorem 1.
\begin{theorem}
If $\omega$ is antipalindromic, the corresponding $\phi$ has the palindromic property.
\end{theorem}
\begin{proof} In Definition 1, we may substitute $\phi_\star = 0, \phi^\star = \delta = \zeta$. Hence it is enough to show that \[\forall j,\ 0 \le j \le m\zeta: \pi_m(j) = \pi_m(m \zeta - j).\]
$\pi_m(m \zeta - j) = \lambda(\phi^{(m)}(x)=m\zeta-j)=
\\=\sum\limits_{(\phi_1, \ldots, \phi_m)} \lambda\big(\phi(x)=\phi_1, \phi^{(2)}(x) = \phi_2, \ldots, \phi^{(m)}(x)=\phi_m\big) \mathbb{I} (\phi_1 + \ldots + \phi_k = m\zeta - j) = 
\\ = \sum\limits_{(\phi_1, \ldots, \phi_m)} \lambda\big(\phi(x)=\phi_1, \phi^{(2)}(x) = \phi_2, \ldots, \phi^{(m)}(x)=\phi_m\big) \mathbb{I} \big((\zeta-\phi_1) + \ldots + (\zeta-\phi_m) = j\big)$\\
Using \Cref{distLemma}, this equals to:\\
$\sum\limits_{(\phi_1, \ldots, \phi_m)} \lambda\big(\phi(x)=\zeta-\phi_m, \phi^{(2)}(x) = \zeta-\phi_{m-1}, \ldots, \phi^{(m)}(x)=\zeta-\phi_1\big) \mathbb{I} \big(\sum\limits_k (\zeta-\phi_k) = j\big) = \left< \psi_k := \zeta - \phi_k \right> = 
\\ =\sum\limits_{(\psi_1, \ldots, \psi_m)} \lambda(\phi(x)=\psi_1, \phi^{(2)}(x) = \psi_2, \ldots, \phi^{(m)}(x)=\psi_m) \mathbb{I} (\psi_1 + \ldots + \psi_m = j) =  \lambda(\phi^{(m)}(x)=j)= \\ =\pi_m(j)$
\end{proof}
We may describe a larger set of functionals $\phi$ having the palindromic property with the use of following lemma.
\begin{lemma}[On the affine transformations]\label{affineLemma}
Let $\omega:[0,p-2] \rightarrow [0, \zeta]$ be antipalindromic, then for any $a > 0, b \ge 0$:\\
$\phi'(x) = \begin{cases}
                    a \omega(x_0) + b, & 0 \le x_0 \le p - 2 \\
                    \phi'(\sigma x), & x_0 = p - 1
                \end{cases}$ is antipalindromic.
\end{lemma}
\begin{proof}
Let $\pi_m'(j) = \lambda(\phi'^{(m)}(x)=j)$. In terms of Definition 1, $\phi_\star = b, \phi^\star = a\zeta + b, \delta = a\zeta$. Let us prove that $\pi_m'(mb + j) = \pi_m' (m(a\zeta+b)-j)$.\\
First, we perform the division with remainder: $j = qa + r$. It follows from the construction of $\phi'$ that $\pi_m'(mb + qa + r) = 0$ if $r \ne 0$. Yet, $m(a\zeta+b)-j = m(a\zeta+b)-qa - r = mb + (m\zeta -q)a - r$ and hence $\pi_m' (m(a\zeta+b)-j) = 0$ if $r \ne 0$.\\
Thus, it remains to prove that $\pi_m'(mb + qa) = \pi_m' (m(a\zeta+b)-qa)$.
Note that we may restore the values of $\phi$ produced by $\omega$ from the values of $\phi'$. Indeed, consider the bijection $i: \{b, a + b, 2a + b \ldots, a\zeta + b\} \rightarrow [0, \zeta]$ such that $i(j) = \frac{j-b}{a}$. It's easy to see that $i(\phi'(x))=\phi(x)$. Subsequently, we may define $i^{(m)}(j) = \frac{j-mb}{m}$ and conclude $i^{(m)}(\phi'^{(m)}(x))=\phi(x)$.\\
Now let us prove $\pi_m'(mb + qa) = \pi_m' (m(a\zeta+b)-qa)$ using the fact that $i^{(m)}$ is bijective.\\
\[\pi_m'(mb + qa) = \lambda(\phi'^{(m)}(x)=mb + qa) = \lambda\big(i^{(m)}(\phi'^{(m)}(x))=i^{(m)}(mb + qa)\big)  =\]\[= \lambda(\phi^{(m)}(x)= \frac{mb-qa-mb}{m})= \lambda(\phi^{(m)}(x)= q)=\pi_m(q)\]
Similarly, $\pi_m' (m(a\zeta+b)-qa) = \pi_m(m\zeta - q)$. Since $\omega$ is antipalindromic, it follows from Theorem 2 that $\pi_m(q) = \pi_m(m \zeta - q)$ and then $\pi_m'(mb + qa) = \pi_m' (m(a\zeta+b)-qa)$.
\end{proof}
\begin{proposition}[On inheritance of polindromic property]
Let $\phi$ have the polindromic property, and let there be $\phi'$ such that for any $m \in \mathbb{N}$ there exists bijection $i^{(m)}: \mathrm{Ran}\ \phi'^{(m)} \rightarrow \mathrm {Ran}\ \phi^{(m)}$. Then $\phi'$ has the polindromic property.
\end{proposition}
The proof of this must be the same as for \Cref{affineLemma}.

\section{Recurrence formulae for linear $\omega$}
\begin{notation} We denote $n$th triangle number $\frac{n(n+1)}{2}$ as $\Delta_n$ \end{notation}
\begin{lemma}\label{simpleCase} $P_{pm}^p(t)=t^{m\Delta_{p-2}}P_m^p(t)$
\end{lemma}
\begin{proof}
Let $x \in \Gamma'$. Recalling the structure of $\{\phi(S^j x)\}_{j \ge 0}$, the value of $\phi^{(pm)}(x)$ is the sum of:
\begin{itemize}
\item the order-$0$ points. There are exactly $(p-1)m$ points following the repeating pattern $\ldots,(p-2),0,1,2,...,(p-2),0,1,\ldots$. Their contribution is $m$ times the sum of integers $0,1,\ldots,(p-2)$ which is $\frac{(p-1)(p-2)}{2}m=m\Delta_{p-2}$
\item the higher-order points. Their contribution is $\phi^{(m)}(\sigma x)$.
\end{itemize}
Hence $\phi^{pm}(x) = m\Delta_{p-2} + \phi^{(m)}(\sigma x)$. By the definition of polynomials $P_m(t)$ it implies 
$P_pm(t) = \mathbb{E}_\lambda\left[ t^{\phi^{(pm)}(x)}\right] = 
\mathbb{E}_\lambda\left[ t^{m\Delta_{p-2} + \phi^{(m)}(\sigma x)}\right] = 
t^{m\Delta_{p-2}} \mathbb{E}_\lambda \left[ t^{\phi^{(m)} (\sigma x)}  \right] = 
t^{m\Delta_{p-2} }P_m(t)$
\end{proof}
\begin{notation}
Similarly to $\phi^{(m)}$, we denote $\omega^{(m)}(j) := \omega(j) + \omega(j+1) + \ldots + \omega(j+m-1)$ if $j + m < p - 1$.
\end{notation}
\begin{proposition} $\omega^{(k)}(j) = k j + \Delta_{k-1}$
\end{proposition}
\begin{proof}
The value of $\omega^{(k)}(x_0)$ is the sum of an arithmetic progression $x_0, x_0 + 1, \ldots, x_0 + k-1$.
\end{proof}
\begin{lemma} \label{phiLemma}
Let $0 < k < p$. Consider $x=(x_0,x_1,\ldots) \in \Gamma'$.

$\phi^{(pm+k)}(x)= 
\begin{cases}
\omega^{(k)}(x_0) + m\Delta_{p-2} + \phi^{(m)}(\sigma x), & x_0 < p-k\\
\omega^{(p-1-x_0)}(x_0) + \Delta_{x_0+k-p-1}+m\Delta_{p-2}+\phi^{(m+1)}(\sigma x), & x_0 \ge p-k
\end{cases}$
\end{lemma}
\begin{proof}
First, by the definition of $\phi^(k)$ we state $$\phi^{(pm+k)}(x) = \phi(x) + \phi(Sx) + \ldots + \phi(S^{k-1}x) + \phi^{(pm)}(S^k x)= \phi^{(k)}(x) + \phi^{(pm)}(S^k x).$$
With the use of the previous lemma holds the equality $\phi^{(pm+k)}(x) =  \phi^{(k)}(x) + m\Delta_{p-2} + \phi^{(m)}(\sigma S^k x)$.
By the definition of $\phi(x)$ there are two cases in the computation of $\phi^{(pm+k)}(x)$:
\begin{itemize}
\item The regular case $x_0 < p-k $. In this case each term in $\phi^{(k)}(x)$ is computed directly: $\phi^{(k)}(x) = \omega^{(k)}(x)$. Yet $\sigma S^k x = \sigma x$: since $S^k$ affects only the first digit of $x$, its effect if erased from $\sigma S^k x$. We may compute $\phi^{(pm+k)}(x)=\omega^{(k)}(x_0)+m\Delta_{p-2}+\phi^{(m)}(\sigma x)$.
\item In another case, if $x_0 \ge p-k $, some term  in $\phi^{(k)}(x)$ evaluates with recursion: there exists $j \in [0,k-1]$ such that $S^j x$ begins with $p-1$ and hence $\phi(S^j x) = \phi(\sigma x)$. Therefore we may not compute $\phi^{(k)}(x)$ directly. Instead, we divide it into $\phi^{(p-1-x_0)}(x)=\omega^{(p-1-x_0)}(x)$, which is computed directly, and the rest of the terms. The latter form the sum $Z := \phi(S^j x) + \phi(S^{j+1}x) + \ldots + \phi(S^{k-1})$. Consider this sum. As stated before, $S^j x$ begins with $p-1$, so the first digits of $S^{j+1}x, S^{j+2}x, \ldots, S^{k-1}$ are $0,1,\ldots, x_0+k-p-1$. We know $\phi(S^j x) = \phi(\sigma x)$, and, knowing the first digits of the following terms, we may evaluate $Z = \phi(\sigma x) + 0 + 1 + \ldots + (x_0+k-p-1) =  \phi(\sigma x) + \Delta_{x_0+k-p-1}$. Now we aggregate $\phi^{(k)}(x)=\omega^{(p-1-x_0)}(x) + \Delta_{x_0+k-p-1}+\phi(\sigma x)$. Using the equality $\phi^{(m)}(\sigma S^k x) = \phi^{(m)}(S\sigma x) = \phi^{(m+1)}(\sigma x) - \phi(\sigma x)$, we get 
$$\phi^{(pm+k)}(x) =  \phi^{(k)}(x) + m\Delta_{p-2} + \phi^{(m)}(\sigma S^k x) = $$ $$ = \omega^{(p-1-x_0)}(x) + \Delta_{x_0+k-p-1}+\phi(\sigma x)+ m\Delta_{p-2} + \phi^{(m+1)}(\sigma x) - \phi(\sigma x) = $$ $$ = \omega^{(p-1-x_0)}(x_0) + \Delta_{x_0+k-p-1}+m\Delta_{p-2}+\phi^{(m+1)}(\sigma x).$$
\end{itemize}
\end{proof}
\begin{lemma} \label{complexCase}
Let $0 < k < p$.

$$P_{pm+k}^p(t)  =   \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-1}}\sum\limits_{j=0}^{p-k-1} t^{jk} P_m^p(t) 
      + \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-2}}\sum\limits_{j=0}^{k-1} t^{j(p-k)} P_{m+1}^p(t) $$
\end{lemma}
\begin{proof}
This lemma is shown using the previous result and the law of total expectation.
$$ P_{pm+k}(t)= \mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)}\right] = $$
$$= \mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)}| x_0 < p-k \right]\lambda(x_0<p-k) + \mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)} | x_0 \ge p-k \right]\lambda( x_0 \ge p-k )$$
\begin{equation}\label{eq:totalExp}= \mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)}| x_0 < p-k \right]\frac{p-k}{p} + \mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)} | x_0 \ge p-k \right]\frac{k}{p}
\end{equation}
Now we calculate each conditional expectations according to \Cref{phiLemma}. Note that the digits of $x$ are mutually independent, hence $x_0 \indep (x_1, x_2,\ldots)$ and then $x_0 \indep \sigma x$, which is used in the evaluation.
$$\mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)}| x_0 < p-k \right] = 
\mathbb{E}_\lambda\left[ t^{ \omega^{(k)}(x_0) + m\Delta_{p-2} + \phi^{(m)}(\sigma x) }| x_0 < p-k \right] = $$ 
$$ =t^{m\Delta_{p-2}}
\mathbb{E}_\lambda\left[ t^{ \omega^{(k)}(x_0)} | x_0 < p-k \right]
 \mathbb{E}_\lambda\left[ t^{\phi^{(m)}(\sigma x) } \right]
$$
The distributions of $\sigma x$ and $x$ are the same, which implies
$\mathbb{E}_\lambda\left[ t^{\phi^{(m)}(\sigma x) } \right] = \mathbb{E}_\lambda\left[ t^{\phi^{(m)}(x) }\right] = P_m^p(t)$. It remains to evaluate $\mathbb{E}_\lambda\left[ t^{ \omega^{(k)}(x_0) | x_0 < p-k }\right]$:
$$ \mathbb{E}_\lambda\left[ t^{ \omega^{(k)}(x_0)} | x_0 < p-k \right] =
 \frac{1}{p-k}\sum\limits_{j=0}^{p-k-1} t^{ \omega^{(k)}(j)} = 
 \frac{1}{p-k}\sum\limits_{j=0}^{p-k-1} t^{kj + \Delta_{k-1}} =
 \frac{1}{p-k} t^{\Delta_{k-1}}\sum\limits_{j=0}^{p-k-1} t^{kj}$$
 Thus, \begin{equation}\label{eq:firstCondExp}\mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)}| x_0 < p-k \right] =
  \frac{1}{p-k} t^{m\Delta_{p-2} + \Delta_{k-1}}\sum\limits_{j=0}^{p-k-1} t^{jk} P_m^p(t)
   \end{equation}

   Similarly, we evaluate the second conditional expectation. 
 $$\mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)}| x_0 \ge p-k \right] = 
 \mathbb{E}_\lambda\left[ t^{\omega^{(p-1-x_0)}(x_0) + \Delta_{x_0+k-p-1}+m\Delta_{p-2}+\phi^{(m+1)}(\sigma x)}| x_0 \ge p-k \right] = $$
 $$ =t^{m\Delta_{p-2}}
 \mathbb{E}_\lambda\left[ t^{\omega^{(p-1-x_0)}(x_0) + \Delta_{x_0+k-p-1}} | x_0 \ge p-k \right] 
 \mathbb{E}_\lambda\left[ t^{\phi^{(m+1)}(\sigma x) } \right] = $$
  $$ =t^{m\Delta_{p-2}}
    P_{m+1}^p(t)
 \mathbb{E}_\lambda\left[ t^{\omega^{(p-1-x_0)}(x_0) + \Delta_{x_0+k-p-1}} | x_0 \ge p-k \right] = $$
  $$ = \frac{1}{k} t^{m\Delta_{p-2}} P_{m+1}^p(t)
  \sum\limits_{j=p-k}^{p-1} t^{  (p-1-j) j + \Delta_{p-2-j} + \Delta_{j+k-p-1}}  = $$

We substitute the index $j$ with $q = p-1-j$ and simplify the exponent:
$$  (p-1-j) j + \Delta_{p-2-j} + \Delta_{j+k-p-1} =
    q(p-1-q) + \Delta_{q-1} + \Delta_{k-2-q}=
   q(p-k)+\Delta_{k-2} 
$$   
Now we collect the result: 
\begin{equation}\label{eq:secondCondExp}
\mathbb{E}_\lambda\left[ t^{\phi^{(pm+k)}(x)}| x_0 \ge p-k \right] 
 =  \frac{1}{k} t^{m\Delta_{p-2} + \Delta_{k-2}} P_{m+1}^p(t)
  \sum\limits_{q=0}^{k-1} t^{q(p-k)}
  \end{equation}
It remains to substitute (\ref{eq:firstCondExp}) and (\ref{eq:secondCondExp}) into (\ref{eq:totalExp}) to obtain the recurrent equation:
$$P_{pm+k}(t)  =   \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-1}}\sum\limits_{j=0}^{p-k-1} t^{jk} P_m^p(t) 
      + \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-2}}\sum\limits_{j=0}^{k-1} t^{j(p-k)} P_{m+1}^p(t) $$
\end{proof}
Bringing together the results of \Cref{simpleCase} and \Cref{complexCase}, we obtain the general recurrent law for $P_{n}^p(t)$
\begin{theorem}[Recurrent formulae on $P_{n}^p(t)$] 
\begin{align}
\label{P_pm} P_{pm}^p(t) & =  t^{m\Delta_{p-2}}\ P_m^p(t) \\
\label{P_pm+k} P_{pm+k}^p(t) & =  \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-1}}\sum\limits_{j=0}^{p-k-1} t^{jk} P_m^p(t)  + \frac{1}{p} t^{m\Delta_{p-2} + \Delta_{k-2}}\sum\limits_{j=0}^{k-1} t^{j(p-k)} P_{m+1}^p(t), \ 0 < k < p
 \end{align}
\end{theorem}

\section{Degrees of polynomials}
\subsection{Sequence $\hat{s}_j$}
Here we define the substitution-produced sequence $\{\hat{s}_j\}_{j \ge 0}$. It will be useful to evaluate the degrees of $P_m^p(t)$. We build this sequence recursively. First, fix the parameter value $p \ge 3$.
\begin{definition}[Zero-tier construction]
Let $\hat{s}_0^{(0)} = p-2,\ \hat{s}_j^{(0)} = p-2-(j-1)$ for $j = \overline{1, p-1}$.
\end{definition}
Literally, $\{ \hat{s}_j^{(0)}\}$ is $p-2, p-2, p-3, \ldots, 2, 1, 0$.
\begin{definition} [Number substitution rule]
We substitute
\begin{itemize}
 \item $0$ with $p-2, p-3, \ldots, 2, 1, 0, 0$;
 \item $1$ with $p-2, p-3, \ldots, 2, 1, 1, 0$;
 \item $2$ with $p-2, p-3, \ldots, 2, 2, 1, 0$;
 \item \ldots
 \item $p-3$ with $p-2, p-3, p-3, \ldots, 2, 1, 0$;
 \item $p-2$ with $p-2, p-2, p-3, \ldots, 2, 1, 0$;

 \item etc.
 \end{itemize}
 Each number $q = \overline{0, p-2}$ becomes the descendent sequence of numbers from $p-2$ to $0$ with $q$ taken twice in a row.
\end{definition}
\begin{definition}[First-tier construction]\label{tier1}
 We have built this sequence $\{ \hat{s}_j^{(0)}\}$  of $p$ integers. To define 
$\hat{s}_j^{(1)}$, we substitute each number in $\{ \hat{s}_j^{(0)}\}$ according to the rule. As a result, we have the sequence $\{\hat{s}_j^{(1)}\}$ of $p^2$ integers.
\end{definition}
We continue the procedure ad infinitum, substituting each number in $\{ \hat{s}_j^{(l)}\}$  to define $\{ \hat{s}_j^{(l+1)}\}$ . It is easy to see that
\begin{itemize}
\item on each step the length of the sequence multiplies by $p$ and
\item for each $l \ge 0$,  $\{ \hat{s}_j^{(l)}\}$ is a prefix of $\{ \hat{s}_j^{(l+1)}\}$.
\end{itemize}

There exists a countably infinite sequence $\{ \hat{s}_j^{(\infty)}\}$ which contains each of $\{ \hat{s}_j^{(l)}\}$ as a prefix.
\begin{definition}
Let $\{\hat{s}_j\}_{j \ge 0} := \{ \hat{s}_j^{(\infty)}\}_{j \ge 0}$.
\end{definition}
\begin{example}
Let $p=4$.

$\begin{array}{|c|cccc|cccc|cccc|cccc|}
\hline
\{ \mathbf{\hat{s}_j^{(0)}}\} & 2 & & & & 2 & & & & 1 & & & & 0 & & & \\
\hline
\{ \mathbf{\hat{s}_j^{(1)}}\} & 2 & 2 & 1 & 0 & 2 & 2 & 1 & 0 & 2 & 1 & 1 & 0 & 2 & 1 & 0 & 0 \\
\hline
\{ \mathbf{\hat{s}_j^{(2)}}\} &
\text{\small{2210}} & \text{\small{2210}} & \text{\small{2110}} & \text{\small{2100}} &
\text{\small{2210}} & \text{\small{2210}} & \text{\small{2110}} & \text{\small{2100}} &
\text{\small{2210}} & \text{\small{2110}} & \text{\small{2110}} & \text{\small{2100}} &
\text{\small{2210}} & \text{\small{2110}} & \text{\small{2100}} & \text{\small{2100}} \\
\hline
 \mathbf{\ldots} & & & & & & & & & & & & & & & & \\
\hline
\end{array} $
\end{example}

\begin{proposition}[Characteristic property of $\{\hat{s}_j\}$]\label{charProp}
For each $m > 0,\ 0 \le j \le p-1$: $\hat{s}_{pm+j}=\hat{s}_{p(p-1-s_m) + j}$
\end{proposition}
\begin{proof}
First, we find the number of the first tier where the $(pm+j)$-th element appears: $l = \log_p{m}$. This $l$-th tier consists of $p$ blocks, each having $p^l$ elements, and each block represents an element from tier $l-1$. The element  $\hat{s}_{pm+j}$ belongs to the $m$-th block of tier $l$, which is the expansion of the $m$-th element in  tier $l-1$. Hence,  
$\hat{s}_{pm+j}$ is the $j$-th element in the substitution of $\hat{s}_m$.

Note that all the substitutions are listed in tier $1$, which consists of the elements from $\hat{s}_0$ to $\hat{s}_{p^2-1}$.
Rigorously, the substitution of $q$ if represented with the items from $\hat{s}_{p + p(p-2-q)}$ to $\hat{s}_{p + p(p-2-q) + p-1}$. If we desire the $j$-th element of this substitution (counting from 0), it is $\hat{s}_{p + p(p-2-q) + j}$.

Using this result and simplifying the index, we show that the $j$-th element in the substitution of $\hat{s}_m$ is $\hat{s}_{p(p-1-s_m) + j}$. 
\end{proof}
\begin{example}
Let $p=4$. Our goal is to find $\hat{s}_{141}$.
$$ \hat{s}_{141} = \hat{s}_{4 \cdot 35 + 1} = \hat{s}_{4(3-\hat{s}_{35}) +1}$$
$$ \hat{s}_{35} = \hat{s}_{4 \cdot 8 + 3} = \hat{s}_{4(3-s_{8}) +3}$$
$$\hat{s}_8 = 2 \implies \hat{s}_{35}=\hat{s}_{4(3-2)+3}=\hat{s}_7=0$$
$$\hat{s}_{141}=\hat{s}_{4(3-\hat{s}_{35}) +1} = \hat{s}_{4(3-0)+1} = \hat{s}_{13}=1$$

The answer is 1.

\end{example}

With the use of this property, we may make a more handy definition of $\hat{s}$. Indeed, as shown in the example, for any $j \ge p^2$ we may find $\hat{s}_j$ applying the equation from \Cref{charProp} sufficient number of times. Hence it is enough to define only the tier 1 (i.e. the values of $\hat{s}_0, \ldots, \hat{s}_{p^2-1}$ to construct the whole sequence.
\begin{definition}\label{sm_good_def}
Define $\{\hat{s}_j^{(1)}\}$ as in \Cref{tier1}. Let
$$ \hat{s}_{pm+j} =
 \begin{cases}
     \hat{s}_{pm+j}^{(1)}, & m < p \\
     \hat{s}_{p(p-1-s_m) + j}, & m \ge p
 \end{cases}
$$
\end{definition}

We will use this definition further.

\subsection{Recurrent formulae for degrees}
\begin{definition}
Let $m \ge 0$. We define $d_m^p := \mathrm{deg}\ P_m^p$.
\end{definition}
\begin{proposition} \label{dm_eq}
For any $m \ge 0,\ 0 < k < p$:
\begin{align*}
  d_{pm}^p &= m \Delta_{p-2} + d_m^p \\
  d_{pm+k}^p &= m \Delta_{p-2}+(p-k)(k-1)+\Delta_{k-2}+d_m^p + \mathrm{max}\{p-k-1, d_{m+1}^p-d_m^p\}
\end{align*}
\end{proposition}
\begin{proof}
Directly from the (\ref{P_pm}) and (\ref{P_pm+k}) we obtain the equations for degrees:
\begin{align*}
  d_{pm}^p &= m \Delta_{p-2} + d_m^p \\
  d_{pm+k}^p &= \mathrm{max}\{ m\Delta_{p-2} + \Delta_{k-1} + k(p-k-1) + d_m^p,\     
                                m\Delta_{p-2} + \Delta_{k-2} + (p-k)(k-1) + d_{m+1}^p \}
\end{align*}
To show this proposition, it is enough to take the common terms out of the maximum operator.
\end{proof}
The following holds for aribtary fixed $p \ge 3$. Keeping this in mind, we use $d_j$ instead of $d_j^p$ for short.
\begin{definition} \label{sm_def}
For any $m \ge 0$, we define $s_m := d_{m+1}^p - d_m^p $.
\end{definition}
Our goal is to show that this sequence $\{s_m\}_{m \ge 0}$ is equal to $\{\hat{s}_m\}$ defined in the previous Subsection. First, we deduce recurrent equations on $s_m$.
\begin{proposition}
\begin{equation} \label{sm_eq}
s_{pm+k} = p-k-1 - \mathbbm{1}\{s_m < p-1-k\}
\end{equation}
\end{proposition}
\begin{proof}
We will show this using \Cref{dm_eq} and \Cref{sm_def}. If we try to evaluate $s_j$, it collapses into two cases:
\begin{itemize}
\item Let $k<p-1 $. $$s_{pm+k} = d_{pm+k+1}-d_{pm+k} =$$ $$= m\Delta_{p-2} + (p-k-1)k + \Delta_{k-1} + d_m + \max\{p-k-1, d_{m+1}-d_m\} - $$ $$ - m\Delta_{p-2}-(p-k)(k-1)-\Delta_{k-2}-d_m-\max\{p-k,d_{m+1}-d_m\}=$$ $$ =p-k-1+\max\{p-k-1, d_{m+1}-d_m\}-\max\{p-k,d_{m+1}-d_m\};$$
We substitute $d_{m+1}-d_m = s_m$ and observe that $$\max\{p-k-1, d_{m+1}-d_m\}-\max\{p-k,d_{m+1}-d_m\}= \mathbbm{1}\{s_m < 1-k\}.$$ This subsitution leads to exaclty the equation in the proposition.
\item $s_{pm+p-1} = d_{p(m+1)}-d_{pm+3} = \\=(m+1)\Delta_{p-2}+d_{m+1}-m\Delta_{p-2}-(p-(p-1))(p-1-1)-\Delta_{p-1-2}-d_m-\max\{0, d_{m+1}-d_m\}=\\=\Delta_{p-2}+(d_{m+1}-d_m)-(p-2)-\Delta_{p-3}-(d_{m+1}-d_m)=0$
\end{itemize}

A straightforward substitution shows that the second case is governed by the same equation.
\end{proof}

\begin{lemma}\label{sm1}
For $j = 0, 1, \ldots, p^2-1:\ s_j=\hat{s}_j^{(1)}$
\end{lemma}
\begin{proof}
We compute the initial $p^2$ terms of this sequence. Recall that $P_0^p(t)=1,\ P_1^p(t) = \frac{1}{p-1} \sum\limits_{j=0}^{p-2} t^j$, hence $d_0 = 0,\ d_1 = p-2$ and $s_0 = p-2$. We will use this as the initial condition in the recurrent equation from (\ref{sm_eq}).
Observe the equation (\ref{sm_eq}). The right part consists of a periodic term $p-k-1$ with period $p$ and the step function $-\mathbbm{1}\{s_m < p-1-k\}$ which is determined by the number of period. Thus we evaluate the sequence $\{s_j\}$ period-wise.

In the first period, $\{s_{0p+k}\}$, we have already shown $s_0=p-2$. The step function is $-\mathbbm{1}\{k < 1\}$, so it affects none of the terms $s_1, \ldots, s_{p-1}$. They form the descending progression $p-2, p-3, \ldots, 1, 0$. Now we have the first $p$ terms of $\{s_j\}$ which are to be used to obtain the subsequent terms.

In the second period, the step function is $-\mathbbm{1}\{s_1 < p-1-k\}$. We know that $s_1=2$, thus the second period repeats the first. In the third period we have $-\mathbbm{1}\{s_1 < p-1-k\}=-\mathbbm{1}\{p-3 < p-1-k\} = -\mathbbm{1}\{k < 2\}$ which affects first two terms. The third period is $p-2,p-3,p-3,p-4, \ldots, 1, 0$.

Similarly, we use all of the terms in the first period to evaluate each of $s_j$ for $j < p^2$. 
It is shown with ease that they the same as in $\hat{s}_j^{(1)}$
\end{proof}

\begin{proposition} \label{sm_estimate}
$\forall j \ge 0:\ s_j \in \{0, 1, \ldots, p-2\}$
\end{proposition}
\begin{proof}
The explicit evaluation of $s_0, s_1, \ldots, s_{p-1}$ shows 
\begin{equation}\label{subset}
\{0,1,\ldots,p-2\} \subset \mathrm{Ran\ } \{s_j\}
\end{equation}

Obviously, $s_j \ge 0$.
We prove by induction that $\forall m \ge p,\ 0 < k < p-1:\ s_{pm+k} < p-1$. The order $\log_p m =: u $ is the parameter of induction. The explicit evaluation of $s_p, s_{p+1}, \ldots, s_{p^2-1}$ shows the base of induction. Let the statement be true for an arbitary $u$. When computing the values of $s_j$ for order $u+1$ from (\ref{sm_eq}), we refer to the values of $s_m$ of order $u$. In (\ref{sm_eq}), $p-k-1 \le p-1$ and $- \mathbbm{1}\{s_m < p-1-k\} \le 0$, thus $s_j \le p-1$. Let us show that the equality is never reached. Indeed, $p-k-1 = p-1$ if true only for $k=0$. When we substitute $k=0$ into the step function, we get $- \mathbbm{1}\{s_m < p-1\}$. For $s_m$ of order $u$, the statement in brackets is true due to the induction hypothesis. So, in order $u+1$ the statement $s_j < p-1$ holds. The inductive step is now shown.

Now we may assert $\mathrm{Ran}\ \{s_j\} \subset \{0,1,\ldots,p-2\}$. Bringing together this and (\ref{subset}), we infer the proposition.
\end{proof}
\begin{lemma} \label{sm2}
For each $m \ge p,\ 0 < k < p-1:\ s_{pm+k}=s_{p(p-1-s_m)+k}$
\end{lemma}
\begin{proof}
First, we observe that for any $j \ge 0$, $s_j = s_{p-1 - s_j}$. Using the \Cref{sm_estimate}, it is enough to show the statement for $s_j=0,1,\ldots,p-2$. For example, let $s_j=0$. We make sure that $0 = s_{p-1}$ which we know from the direct computation. Similarly, we do this for any $s_j=1,\ldots,p-2$.

To prove the lemma, we will this equation. Let $m \ge p,\ 0 < k < p-1$. From (\ref{sm_eq}) we obtain $s_{pm+k}=p-k-1 - \mathbbm{1}\{s_m < p-1-k\}$ and $s_{p(p-1-s_m)+k}= p-k-1 - \mathbbm{1}\{s_{p-1-s_m} < p-1-k\}$. One may easily observe that the equality $s_m=p-1-s_m$ turns this into an identity.
\end{proof}
\begin{theorem}
For each $j \ge 0$: $s_j=\hat{s}_j$.
\end{theorem}
\begin{proof}
The results of \Cref{sm1} and \Cref{sm2} show that \Cref{sm_good_def} is true for $\{s_j\}_{j \ge 0}$.
\end{proof}
\begin{corollary}
$d_m = \sum\limits_{j=0}^{m-1} \hat{s}_j$
\end{corollary}
\end{document}