### 2018-06-22 presentations/seminar/title.tex
presentations

%\title[On new examples of mixing local rank one transformations] % (optional, use only with long paper titles)
%{New examples of mixing local rank one transformations}

\title[Chacon polynomials] % (optional, use only with long paper titles)
{Generalized Chacon polynomial constructions}

%\subtitle
%{Include Only If Paper Has a Subtitle}

\author%[Author, Another] % (optional, use only with lots of authors)
{V.V.~Slyusarev}
% - Give the names in the same order as the appear in the paper.
% - Use the \inst{?} command only if the authors have different
%   affiliation.

\institute[https://arxiv.org/abs/1802.09372] % (optional, but mostly needed)
{
  %\inst{1}%
  %AMU / MIPT
  Moscow institute of physics and technology $\quad$ %\\ 
  % \includegraphics[height=6.3mm]{amu_logo_xsmall.png}
}
% - Use the \inst command only if there are several affiliations.
% - Keep it simple, no one is interested in your street address.

\date[ 25.11.2017] % (optional, should be abbreviation of conference name)
{\normalsize https://arxiv.org/abs/1802.09372}
% - Either use conference name or its abbreviation.
% - Not really informative to the audience, more for people (including
%   yourself) who are reading the slides online

\subject{https://arxiv.org/abs/1802.09372}
% This is only inserted into the PDF information catalog. Can be left
% out.